---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# covidmonitor

<!-- badges: start -->
<!-- badges: end -->

The goal of {covidmonitor} is to provide functions for simplifying data merging 
and cleaning for WHO AFRO COVID-19 linelists and monitoring & evaluation data. 
The package further contains templates for automating report production. 

## Installation

Currently the package is not on CRAN. Once it is - you can install the released 
version of epichecks from [CRAN](https://CRAN.R-project.org) with:

``` r
# install.packages("covidmonitor")
```

In order to install the package you will first need to install an extra bit of 
software called [Rtools](https://cran.r-project.org/bin/windows/Rtools/).  
You can download the installer from: https://cran.r-project.org/bin/windows/Rtools/  
Please install the version highlighted in green. 

Once this is installed and you have restarted your computer, 
the development version of the package can be installed from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("R4IDSR/covidmonitor")
```

If you do not have a LaTeX editor installed on your computer, then please run the
following code to install TinyTex. This is needed in order to be able to create
Word and PDF documents using *R markdown*. 

```r
install.packages('tinytex')
tinytex::install_tinytex()
# to uninstall TinyTeX, run tinytex::uninstall_tinytex() 
```


## Country linelists

### Folder structure 

Functions and templates are set up to run based on files contained in the 
AFRO WHE HIM - Country LineList sharepoint folder. This folder contains all the 
latest country linelist files which are automatically sorted from files emailed 
to [afrooutbreak@who.int](afrooutbreak@who.int) by countries, using a .net script. 

To ensure that the data pipeline works in an automated fashion, please always email 
the country files to the email address specified. This way the latest file can be
efficiently identified. 

### File naming and format

In order for the *R* functions to run, the files must be named appropriately, with 
the start of the file name containing the three letter ISO code for the country 
followed by a dot (.) and then the date of report; after this the rest of the 
name can be anything. 
For example a file submitted by Namibia on the 29th of January 2021 should have 
"NAM.2021-01-29" at the beginning of the name, and so the full file name could be
"NAM.2021-01-29.Linelist.Namibia COVID-19 Confirmed.xlsx". 

All files must be stored as an excel file (i.e. **xlsx**), those stored as binary 
workbooks (**xlsb**) will throw an error and must be re-saved as **xlsx**. 

**n.b** some countries (e.g. TCD) have saved their date variables in excel using 
the 1904 date system (i.e. date origin is 1st January 1904 rather than 1900), this 
needs to be restructured manually in the excel file otherwise the *R* function 
will produce dates that are off by four years. See [here](https://docs.microsoft.com/en-us/office/troubleshoot/excel/1900-and-1904-date-system) 
for details. 


### Dictionaries 

In order for the functions to work, several data dictionaries are required. 
The first is called [linelist_dictionary.xlsx](https://github.com/R4IDSR/covidmonitor/raw/master/inst/extdata/linelist_dictionary.xlsx) and contains information for 
defining and naming variables, as well as defining which sheet to read in from 
the various country linelists. 
The second is called [cleaning_dictionary.xlsx](https://github.com/R4IDSR/covidmonitor/raw/master/inst/extdata/cleaning_dictionary.xlsx) and contains information for 
cleaning and recoding variables. Details of the required sheet names and contents
are in the table below. 

The functions use default to using dictionaries which are pre-defined within the package, 
however you can also edit these dictionaries by downloading them from GitHub (links
provided above), and then specifying the file path within the function call (see
help files for function argument details). **N.b** sheet and variable names should 
not be changed (otherwise the function will break). 


TODO: add details about how to structure dictionaries (table with diff tabs defined) - create an excel file
and put it in the extdata folder (one sheet per dictionary) 

**Table**: Required variables by sheet in the linelist dictionary. 

```{r linelist_dict}



```

**Table**: Required variables by sheet in the cleaning dictionary




### Templates 

TODO: add details about how to use templates







## Monitoring & Evaluation

### Folder structure 

TODO: add details about where data should be and how should be named etc 

### Dictionaries 

TODO: add details about how to structure dictionaries (table with diff tabs and 
how to set up if not using defaults)

### Templates 

TODO: add details about how to use templates

